### 2024-11-19 ASTR540/hw4/Franz_ASTR540_HW4_NB.tex
nb -> pdf for submission

\documentclass[11pt]{article}

    \usepackage[breakable]{tcolorbox}
    \usepackage{parskip} % Stop auto-indenting (to mimic markdown behaviour)

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % Maintain compatibility with old templates. Remove in nbconvert 6.0
    \let\Oldincludegraphics\includegraphics
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionFormat{nocaption}{}
    \captionsetup{format=nocaption,aboveskip=0pt,belowskip=0pt}

    \usepackage{float}
    \floatplacement{figure}{H} % forces figures to be placed at the correct location
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro

    \usepackage{iftex}
    \ifPDFTeX
        \usepackage[T1]{fontenc}
        \IfFileExists{alphabeta.sty}{
              \usepackage{alphabeta}
          }{
              \usepackage[mathletters]{ucs}
              \usepackage[utf8x]{inputenc}
          }
    \else
        \usepackage{fontspec}
        \usepackage{unicode-math}
    \fi

    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics
                         % to support a larger range
    \makeatletter % fix for old versions of grffile with XeLaTeX
    \@ifpackagelater{grffile}{2019/11/01}
    {
      % Do nothing on new versions
    }
    {
      \def\Gread@@xetex#1{%
        \IfFileExists{"\Gin@base".bb}%
        {\Gread@eps{\Gin@base.bb}}%
        {\Gread@@xetex@aux#1}%
      }
    }
    \makeatother
    \usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
    \adjustboxset{max size={0.9\linewidth}{0.9\paperheight}}

    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    % The default LaTeX title has an obnoxious amount of whitespace. By default,
    % titling removes some of it. It also provides customization options.
    \usepackage{titling}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage{array}     % table support for pandoc >= 2.11.3
    \usepackage{calc}      % table minipage width calculation for pandoc >= 2.11.1
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines
    \usepackage{mathrsfs}

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}
    \definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
    \definecolor{ansi-default-inverse-bg}{HTML}{000000}

    % common color for the border for error outputs.
    \definecolor{outerrorbackground}{HTML}{FFDFDF}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatibility definitions
    \def\gt{>}
    \def\lt{<}
    \let\Oldtex\TeX
    \let\Oldlatex\LaTeX
    \renewcommand{\TeX}{\textrm{\Oldtex}}
    \renewcommand{\LaTeX}{\textrm{\Oldlatex}}
    % Document parameters
    % Document title
    \title{Franz\_ASTR540\_HW4\_NB}

% Pygments definitions
\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\@namedef{PY@tok@w}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\@namedef{PY@tok@c}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cp}{\def\PY@tc##1{\textcolor[rgb]{0.61,0.40,0.00}{##1}}}
\@namedef{PY@tok@k}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kt}{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\@namedef{PY@tok@o}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@ow}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\@namedef{PY@tok@nb}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@nf}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@nc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@nn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@ne}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.80,0.25,0.22}{##1}}}
\@namedef{PY@tok@nv}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@no}{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\@namedef{PY@tok@nl}{\def\PY@tc##1{\textcolor[rgb]{0.46,0.46,0.00}{##1}}}
\@namedef{PY@tok@ni}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
\@namedef{PY@tok@na}{\def\PY@tc##1{\textcolor[rgb]{0.41,0.47,0.13}{##1}}}
\@namedef{PY@tok@nt}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@nd}{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\@namedef{PY@tok@s}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sd}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@si}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
\@namedef{PY@tok@se}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.36,0.12}{##1}}}
\@namedef{PY@tok@sr}{\def\PY@tc##1{\textcolor[rgb]{0.64,0.35,0.47}{##1}}}
\@namedef{PY@tok@ss}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@sx}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@m}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@gh}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\@namedef{PY@tok@gu}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\@namedef{PY@tok@gd}{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\@namedef{PY@tok@gi}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.52,0.00}{##1}}}
\@namedef{PY@tok@gr}{\def\PY@tc##1{\textcolor[rgb]{0.89,0.00,0.00}{##1}}}
\@namedef{PY@tok@ge}{\let\PY@it=\textit}
\@namedef{PY@tok@gs}{\let\PY@bf=\textbf}
\@namedef{PY@tok@gp}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\@namedef{PY@tok@go}{\def\PY@tc##1{\textcolor[rgb]{0.44,0.44,0.44}{##1}}}
\@namedef{PY@tok@gt}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\@namedef{PY@tok@err}{\def\PY@bc##1{{\setlength{\fboxsep}{\string -\fboxrule}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}}
\@namedef{PY@tok@kc}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kd}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kn}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@kr}{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@bp}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\@namedef{PY@tok@fm}{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\@namedef{PY@tok@vc}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vg}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vi}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@vm}{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\@namedef{PY@tok@sa}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sb}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sc}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@dl}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@s2}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@sh}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@s1}{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\@namedef{PY@tok@mb}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mf}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mh}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mi}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@il}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@mo}{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\@namedef{PY@tok@ch}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cm}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cpf}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@c1}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}
\@namedef{PY@tok@cs}{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.24,0.48,0.48}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % For linebreaks inside Verbatim environment from package fancyvrb.
    \makeatletter
        \newbox\Wrappedcontinuationbox
        \newbox\Wrappedvisiblespacebox
        \newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}}
        \newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}}
        \newcommand*\Wrappedcontinuationindent {3ex }
        \newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox}
        % Take advantage of the already applied Pygments mark-up to insert
        % potential linebreaks for TeX processing.
        %        {, <, #, %, $, ' and ": go to next line.
        %        _, }, ^, &, >, - and ~: stay at end of broken line.
        % Use of \textquotesingle for straight quote.
        \newcommand*\Wrappedbreaksatspecials {%
            \def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}%
            \def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}%
            \def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}%
            \def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}%
            \def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}%
            \def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}%
            \def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}%
            \def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}%
            \def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}%
            \def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}%
            \def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}%
            \def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}%
            \def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}%
            \def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}%
        }
        % Some characters . , ; ? ! / are not pygmentized.
        % This macro makes them "active" and they will insert potential linebreaks
        \newcommand*\Wrappedbreaksatpunct {%
            \lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}%
            \lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}%
            \lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}%
            \lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}%
            \lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}%
            \lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}%
            \lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}%
            \catcode`\.\active
            \catcode`\,\active
            \catcode`\;\active
            \catcode`\:\active
            \catcode`\?\active
            \catcode`\!\active
            \catcode`\/\active
            \lccode`\~`\~
        }
    \makeatother

    \let\OriginalVerbatim=\Verbatim
    \makeatletter
    \renewcommand{\Verbatim}[1][1]{%
        %\parskip\z@skip
        \sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
        \sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
        \def\FancyVerbFormatLine ##1{\hsize\linewidth
            \vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
                \doublehyphendemerits\z@\finalhyphendemerits\z@
                \strut ##1\strut}%
        }%
        % If the linebreak is at a space, the latter will be displayed as visible
        % space at end of first line, and a continuation symbol starts next line.
        % Stretch/shrink are however usually zero for typewriter font.
        \def\FV@Space {%
            \nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
            \discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
            {\kern\fontdimen2\font}%
        }%

        % Allow breaks at special characters using \PYG... macros.
        \Wrappedbreaksatspecials
        % Breaks at punctuation characters . , ; ? ! and / need catcode=\active
        \OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
    }
    \makeatother

    % Exact colors from NB
    \definecolor{incolor}{HTML}{303F9F}
    \definecolor{outcolor}{HTML}{D84315}
    \definecolor{cellborder}{HTML}{CFCFCF}
    \definecolor{cellbackground}{HTML}{F7F7F7}

    % prompt
    \makeatletter
    \newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
    \makeatother
    \newcommand{\prompt}[4]{
        {\ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}}
    }

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

    \maketitle

    \section{ASTR540 Homework 4}\label{astr540-homework-4}

Noah Franz

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{1}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{import} \PY{n+nn}{warnings}
\PY{k+kn}{from} \PY{n+nn}{functools} \PY{k+kn}{import} \PY{n}{partial}
\PY{k+kn}{from} \PY{n+nn}{copy} \PY{k+kn}{import} \PY{n}{deepcopy}

\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{k+kn}{import} \PY{n+nn}{matplotlib}\PY{n+nn}{.}\PY{n+nn}{pyplot} \PY{k}{as} \PY{n+nn}{plt}

\PY{k+kn}{from} \PY{n+nn}{astropy}\PY{n+nn}{.}\PY{n+nn}{coordinates} \PY{k+kn}{import} \PY{n}{SkyCoord}
\PY{k+kn}{from} \PY{n+nn}{astropy} \PY{k+kn}{import} \PY{n}{units} \PY{k}{as} \PY{n}{u}
\PY{k+kn}{from} \PY{n+nn}{astropy} \PY{k+kn}{import} \PY{n}{constants} \PY{k}{as} \PY{n}{c}
\PY{k+kn}{from} \PY{n+nn}{astropy}\PY{n+nn}{.}\PY{n+nn}{table} \PY{k+kn}{import} \PY{n}{vstack}

\PY{k+kn}{from} \PY{n+nn}{astroquery}\PY{n+nn}{.}\PY{n+nn}{gaia} \PY{k+kn}{import} \PY{n}{Gaia}
\PY{k+kn}{from} \PY{n+nn}{astroquery}\PY{n+nn}{.}\PY{n+nn}{sdss} \PY{k+kn}{import} \PY{n}{SDSS}
\end{Verbatim}
\end{tcolorbox}

    \subsection{Problem 1}\label{problem-1}

Proper Motion of Bo ̈otes III. Let's revisit the proper motion of the
disrupting satellite Bo ̈otes III, as discussed in Carlin \& Sand 2018.
Since that study, there has been the 3rd data release of Gaia (Gaia
DR3), so it is possible that there are more proper motion members in the
catalog, and that the overall uncertainties have improved.

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\item
  First, go to Table 1 of the Carlin paper, which displays all of the
  `radial velocity confirmed' members of Bo ̈otes III, and make your own
  table with updated proper motion values.
\item
  By how much have the uncertainties going from Gaia DR2 to Gaia DR3
  improved? Have there been any large changes in proper motion for any
  of these stars (\textgreater5σ)? \textbar{}
\item
  Calculate your own error-weighted proper motion to Bo ̈otes III,
  compare it with the DR2 value, and make a plot similar to Figure 1 of
  the paper.
\item
  Try to identify further proper motion members in Bo ̈otes III, as
  described in Section 2.3 of the paper. Do the improved measurements of
  Gaia DR3 lead to more plausible proper motion members of Bo ̈otes III?
\end{enumerate}

\subsubsection{Problem 1a}\label{problem-1a}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{2}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} read in the Table 1 data from Carin \PYZam{} Sand (2018)}
\PY{n}{data\PYZus{}dr2} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{read\PYZus{}csv}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Carlin\PYZus{}and\PYZus{}Sand\PYZus{}2018\PYZus{}Table1.txt}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{sep}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}t}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{c+c1}{\PYZsh{} and then clean it up}
\PY{n}{data\PYZus{}dr2} \PY{o}{=} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{dropna}\PY{p}{(}\PY{n}{how}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{all}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}

\PY{n}{data\PYZus{}dr2} \PY{o}{=} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{replace}\PY{p}{(}\PY{p}{\PYZob{}}
    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cdots}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{np}\PY{o}{.}\PY{n}{nan}\PY{p}{,}
    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{N}\PY{l+s+s2}{\PYZdq{}} \PY{p}{:} \PY{k+kc}{False}\PY{p}{,}
    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Y}\PY{l+s+s2}{\PYZdq{}} \PY{p}{:} \PY{k+kc}{True}
\PY{p}{\PYZcb{}}\PY{p}{)}

\PY{n}{data\PYZus{}dr2}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{v\PYZus{}helio\PYZus{}err}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{v}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ +or\PYZhy{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]} \PY{k}{if} \PY{o+ow}{not} \PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{v}\PY{p}{)} \PY{k}{else} \PY{n}{v} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{v\PYZus{}helio}\PY{p}{]}
\PY{n}{data\PYZus{}dr2}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{v\PYZus{}helio}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{v}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ +or\PYZhy{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]} \PY{k}{if} \PY{o+ow}{not} \PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{v}\PY{p}{)} \PY{k}{else} \PY{n}{v} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{v\PYZus{}helio}\PY{p}{]}

\PY{n}{data\PYZus{}dr2}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mu\PYZus{}alpha\PYZus{}err}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{v}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ +or\PYZhy{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}  \PY{k}{if} \PY{o+ow}{not} \PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{v}\PY{p}{)} \PY{k}{else} \PY{n}{v} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{mu\PYZus{}alpha}\PY{p}{]}
\PY{n}{data\PYZus{}dr2}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mu\PYZus{}alpha}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{v}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ +or\PYZhy{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}  \PY{k}{if} \PY{o+ow}{not} \PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{v}\PY{p}{)} \PY{k}{else} \PY{n}{v} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{mu\PYZus{}alpha}\PY{p}{]}

\PY{n}{data\PYZus{}dr2}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mu\PYZus{}delta\PYZus{}err}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{v}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ +or\PYZhy{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}  \PY{k}{if} \PY{o+ow}{not} \PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{v}\PY{p}{)} \PY{k}{else} \PY{n}{v} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{mu\PYZus{}delta}\PY{p}{]}
\PY{n}{data\PYZus{}dr2}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mu\PYZus{}delta}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{v}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ +or\PYZhy{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}  \PY{k}{if} \PY{o+ow}{not} \PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{v}\PY{p}{)} \PY{k}{else} \PY{n}{v} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{mu\PYZus{}delta}\PY{p}{]}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{3}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{data\PYZus{}dr3} \PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{n}{idxs} \PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{n}{radius} \PY{o}{=} \PY{l+m+mi}{1}\PY{o}{*}\PY{n}{u}\PY{o}{.}\PY{n}{arcsec}

\PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{row} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{iterrows}\PY{p}{(}\PY{p}{)}\PY{p}{:}
    \PY{n}{coord} \PY{o}{=} \PY{n}{SkyCoord}\PY{p}{(}\PY{n}{row}\PY{o}{.}\PY{n}{alpha}\PY{p}{,} \PY{n}{row}\PY{o}{.}\PY{n}{delta}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{n}{u}\PY{o}{.}\PY{n}{deg}\PY{p}{)}
    \PY{n}{job} \PY{o}{=} \PY{n}{Gaia}\PY{o}{.}\PY{n}{cone\PYZus{}search}\PY{p}{(}\PY{n}{coord}\PY{p}{,} \PY{n}{radius}\PY{o}{=}\PY{n}{radius}\PY{p}{)}
    \PY{n}{res} \PY{o}{=} \PY{n}{job}\PY{o}{.}\PY{n}{get\PYZus{}results}\PY{p}{(}\PY{p}{)}

    \PY{k}{if} \PY{n+nb}{len}\PY{p}{(}\PY{n}{res}\PY{p}{)} \PY{o}{!=} \PY{l+m+mi}{1}\PY{p}{:}
        \PY{n}{warnings}\PY{o}{.}\PY{n}{warn}\PY{p}{(}\PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Length of result not equal to 1 for }\PY{l+s+si}{\PYZob{}}\PY{n}{row}\PY{o}{.}\PY{n}{source\PYZus{}id}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{!}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{k}{continue}

    \PY{n}{data\PYZus{}dr3}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{res}\PY{p}{)}
    \PY{n}{idxs}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{i}\PY{p}{)} \PY{c+c1}{\PYZsh{} keep track of index for joining}

\PY{n}{data\PYZus{}dr3} \PY{o}{=} \PY{n}{vstack}\PY{p}{(}\PY{n}{data\PYZus{}dr3}\PY{p}{)}\PY{o}{.}\PY{n}{to\PYZus{}pandas}\PY{p}{(}\PY{p}{)}
\PY{n}{data\PYZus{}dr3} \PY{o}{=} \PY{n}{data\PYZus{}dr3}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{idxs}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
/tmp/ipykernel\_466903/39562995.py:11: UserWarning: Length of result not equal to
1 for -9223372036854775808!
  warnings.warn(f"Length of result not equal to 1 for \{row.source\_id\}!")
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{4}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{data\PYZus{}dr3}\PY{o}{.}\PY{n}{columns} \PY{o}{=} \PY{p}{[}\PY{n}{c}\PY{o}{+}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}  \PY{k}{if} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}} \PY{o+ow}{not} \PY{o+ow}{in} \PY{n}{c} \PY{k}{else} \PY{n}{c} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr3}\PY{o}{.}\PY{n}{columns}\PY{p}{]}
\PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{columns} \PY{o}{=} \PY{p}{[}\PY{n}{c}\PY{o}{+}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}dr2}\PY{l+s+s2}{\PYZdq{}} \PY{k}{if} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}dr2}\PY{l+s+s2}{\PYZdq{}} \PY{o+ow}{not} \PY{o+ow}{in} \PY{n}{c} \PY{k}{else} \PY{n}{c} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{data\PYZus{}dr2}\PY{o}{.}\PY{n}{columns}\PY{p}{]}

\PY{n}{data} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{merge}\PY{p}{(}\PY{n}{data\PYZus{}dr2}\PY{p}{,} \PY{n}{data\PYZus{}dr3}\PY{p}{,} \PY{n}{left\PYZus{}index}\PY{o}{=}\PY{k+kc}{True}\PY{p}{,} \PY{n}{right\PYZus{}index}\PY{o}{=}\PY{k+kc}{True}\PY{p}{,} \PY{n}{how}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{outer}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{22}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{get\PYZus{}sdss\PYZus{}colors}\PY{p}{(}\PY{n}{df}\PY{p}{,} \PY{n}{ra\PYZus{}key}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ra}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{dec\PYZus{}key}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{filters}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{u}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{g}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{i}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{z}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{p}{:}

    \PY{n}{all\PYZus{}data} \PY{o}{=} \PY{p}{[}\PY{p}{]}

    \PY{k}{for} \PY{n}{idx}\PY{p}{,} \PY{n}{row} \PY{o+ow}{in} \PY{n}{df}\PY{o}{.}\PY{n}{iterrows}\PY{p}{(}\PY{p}{)}\PY{p}{:}

        \PY{n}{coord} \PY{o}{=} \PY{n}{SkyCoord}\PY{p}{(}\PY{n}{row}\PY{p}{[}\PY{n}{ra\PYZus{}key}\PY{p}{]}\PY{p}{,} \PY{n}{row}\PY{p}{[}\PY{n}{dec\PYZus{}key}\PY{p}{]}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{deg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

        \PY{k}{try}\PY{p}{:}
            \PY{n}{res} \PY{o}{=} \PY{n}{SDSS}\PY{o}{.}\PY{n}{query\PYZus{}region}\PY{p}{(}
                \PY{n}{coord}\PY{p}{,} 
                \PY{n}{radius}\PY{o}{=}\PY{l+m+mi}{1}\PY{o}{*}\PY{n}{u}\PY{o}{.}\PY{n}{arcsec}\PY{p}{,} 
                \PY{n}{spectro}\PY{o}{=}\PY{k+kc}{False}\PY{p}{,}
                \PY{n}{fields}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ra}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{objid}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{+} \PY{n}{filters}\PY{p}{,}
            \PY{p}{)}
        \PY{k}{except} \PY{n+ne}{Exception} \PY{k}{as} \PY{n}{e}\PY{p}{:}
            \PY{n}{all\PYZus{}data}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{row}\PY{p}{)}
            \PY{k}{continue}

        \PY{k}{if} \PY{n}{res} \PY{o+ow}{is} \PY{k+kc}{None} \PY{o+ow}{or} \PY{o+ow}{not} \PY{n+nb}{len}\PY{p}{(}\PY{n}{res}\PY{p}{)}\PY{p}{:}
            \PY{n}{all\PYZus{}data}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{row}\PY{p}{)}
            \PY{n}{warnings}\PY{o}{.}\PY{n}{warn}\PY{p}{(}\PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Skipping }\PY{l+s+si}{\PYZob{}}\PY{n}{row}\PY{o}{.}\PY{n}{DESIGNATION}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ because no SDSS data!}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
            \PY{k}{continue}

        \PY{n}{res} \PY{o}{=} \PY{n}{res}\PY{o}{.}\PY{n}{to\PYZus{}pandas}\PY{p}{(}\PY{p}{)}

        \PY{n}{res\PYZus{}filtered} \PY{o}{=} \PY{n}{res}\PY{p}{[}\PY{n}{res}\PY{o}{.}\PY{n}{g} \PY{o}{\PYZgt{}} \PY{o}{\PYZhy{}}\PY{l+m+mi}{100}\PY{p}{]}

        \PY{k}{if} \PY{o+ow}{not} \PY{n+nb}{len}\PY{p}{(}\PY{n}{res}\PY{p}{)}\PY{p}{:}
            \PY{n}{all\PYZus{}data}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{row}\PY{p}{)}
            \PY{n}{warnings}\PY{o}{.}\PY{n}{warn}\PY{p}{(}\PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Skipping }\PY{l+s+si}{\PYZob{}}\PY{n}{row}\PY{o}{.}\PY{n}{DESIGNATION}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ because no high quality SDSS data!}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
            \PY{k}{continue}

        \PY{c+c1}{\PYZsh{} take the u, g, r, i, z from the closest row }
        \PY{n}{min\PYZus{}sep} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{argmin}\PY{p}{(}
            \PY{n}{coord}\PY{o}{.}\PY{n}{separation}\PY{p}{(}
                \PY{n}{SkyCoord}\PY{p}{(}
                    \PY{n}{res\PYZus{}filtered}\PY{o}{.}\PY{n}{ra}\PY{o}{.}\PY{n}{values}\PY{p}{,} 
                    \PY{n}{res\PYZus{}filtered}\PY{o}{.}\PY{n}{dec}\PY{o}{.}\PY{n}{values}\PY{p}{,} 
                    \PY{n}{unit}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{deg}\PY{l+s+s2}{\PYZdq{}}
                \PY{p}{)}
            \PY{p}{)}
        \PY{p}{)}
        \PY{k}{for} \PY{n}{filt} \PY{o+ow}{in} \PY{n}{filters}\PY{p}{:}
            \PY{n}{row}\PY{p}{[}\PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZob{}}\PY{n}{filt}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZus{}sdss}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{n}{res\PYZus{}filtered}\PY{p}{[}\PY{n}{filt}\PY{p}{]}\PY{o}{.}\PY{n}{values}\PY{p}{[}\PY{n}{min\PYZus{}sep}\PY{p}{]}

        \PY{n}{all\PYZus{}data}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{row}\PY{p}{)}

    \PY{k}{return} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{all\PYZus{}data}\PY{p}{)}

\PY{n}{data} \PY{o}{=} \PY{n}{get\PYZus{}sdss\PYZus{}colors}\PY{p}{(}\PY{n}{data}\PY{p}{,} \PY{n}{ra\PYZus{}key}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ra\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{dec\PYZus{}key}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dec\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{data}\PY{p}{[}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{source\PYZus{}id\PYZus{}dr2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{alpha\PYZus{}dr2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{delta\PYZus{}dr2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mu\PYZus{}alpha\PYZus{}dr2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mu\PYZus{}delta\PYZus{}dr2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{SOURCE\PYZus{}ID\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ra\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dec\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pmra\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pmdec\PYZus{}dr3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{]}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{22}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
          source\_id\_dr2   alpha\_dr2  delta\_dr2 mu\_alpha\_dr2 mu\_delta\_dr2  \textbackslash{}
0   1450811159128205568  208.955653  26.502208        1.078        5.259
1   1450817000283805824  209.013621  26.661679       -0.466       -0.583
2   1450826552290908800  209.128453  26.840015       -1.211       -1.139
3   1450826724089602816  209.159600  26.872888        -1.84       -1.784
4   1451038792395234560  209.174766  26.979976       -1.932        0.579
5   1450823051893369984  209.261163  26.829991       -1.709       -0.332
6   1450803222028673280  209.331958  26.573608       -1.492       -0.928
7   1450833767836315392  209.467272  26.804392       -0.207       -0.927
8   1450828682594980736  209.479318  26.664838       -0.462       -0.479
9   1450835314024556544  209.500251  26.835870       -1.006       -1.389
10  1450748486965605760  209.520145  26.287895       -0.566       -1.793
11  1450842185972291840  209.573497  26.944967       -1.435        0.541
12  1450781163077088000  209.610213  26.651188        0.058       -1.044
13  1450782949783229824  209.668856  26.711657       -2.084       -2.503
14  1450751617997283840  209.681364  26.467320       -5.797       -2.153
15  1450739381635432192  209.700291  26.407019         0.24       -1.969
16  1450785045727265792  209.810712  26.693513       -1.595       -1.315
17  1450788584780752256  209.845448  26.753301       -7.457       -2.497
18  1450783774417361280  209.916855  26.655120      -10.529      -11.886
19 -9223372036854775808  209.489916  26.858467          NaN          NaN
20  1258556500130302080  210.143865  25.931296        -1.34       -0.978

    SOURCE\_ID\_dr3      ra\_dr3    dec\_dr3   pmra\_dr3  pmdec\_dr3
0    1.450811e+18  208.955653  26.502208  -1.957361  -2.901003
1    1.450817e+18  209.013621  26.661679  -1.147453  -0.904130
2    1.450827e+18  209.128453  26.840015  -0.837238  -0.971798
3    1.450827e+18  209.159600  26.872888  -0.857254  -1.206666
4    1.451039e+18  209.174766  26.979976  -1.652480  -0.706780
5    1.450823e+18  209.261163  26.829991  -1.400544  -1.098103
6    1.450803e+18  209.331957  26.573608  -1.378667  -0.721432
7    1.450834e+18  209.467272  26.804392  -1.324451  -0.947451
8    1.450829e+18  209.479318  26.664838  -0.470838  -0.927059
9    1.450835e+18  209.500250  26.835871  -1.611917  -0.534405
10   1.450748e+18  209.520145  26.287895  -0.832462  -1.081806
11   1.450842e+18  209.573497  26.944967  -1.064027  -0.821582
12   1.450781e+18  209.610213  26.651188  -0.600902  -0.834171
13   1.450783e+18  209.668855  26.711657  -3.194013  -1.795397
14   1.450752e+18  209.681363  26.467319  -4.928035  -3.255653
15   1.450739e+18  209.700291  26.407018  -0.215862  -3.008471
16   1.450785e+18  209.810711  26.693513  -1.239291  -1.117548
17   1.450789e+18  209.845446  26.753300  -7.274916  -2.051706
18   1.450784e+18  209.916853  26.655119 -11.452348  -7.981481
19            NaN         NaN        NaN        NaN        NaN
20   1.258557e+18  210.143865  25.931296  -1.235442  -0.889244
\end{Verbatim}
\end{tcolorbox}

    \subsubsection{Problem 1a Discussion}\label{problem-1a-discussion}

A summary of the results of merging the DR2 catalog from Carlin \& Sand
(2018) and the Gaia DR3 measurements of the same objects is shown in the
above table!

\subsubsection{Problem 1b}\label{problem-1b}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{6}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{fig}\PY{p}{,} \PY{n}{axs} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplots}\PY{p}{(}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{n}{figsize}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{18}\PY{p}{)}\PY{p}{)}
\PY{n}{ax1}\PY{p}{,} \PY{n}{ax2}\PY{p}{,} \PY{n}{ax3}\PY{p}{,} \PY{n}{ax4}\PY{p}{,} \PY{n}{ax5}\PY{p}{,} \PY{n}{ax6} \PY{o}{=} \PY{n}{axs}\PY{o}{.}\PY{n}{flatten}\PY{p}{(}\PY{p}{)}

\PY{c+c1}{\PYZsh{} first plot the proper motion from DR2 to DR3}
\PY{n}{ax1}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}
    \PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}dr3}\PY{p}{,} 
    \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}alpha\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{xerr} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}error\PYZus{}dr3}\PY{p}{,}
    \PY{n}{yerr} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}alpha\PYZus{}err\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{none}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{color}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{k}\PY{l+s+s2}{\PYZdq{}}
\PY{p}{)}

\PY{n}{ax1}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha\PYZdl{} (GAIA DR2)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax1}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha\PYZdl{} (GAIA DR3)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{ylim} \PY{o}{=} \PY{n}{ax1}\PY{o}{.}\PY{n}{get\PYZus{}ylim}\PY{p}{(}\PY{p}{)}
\PY{n}{xlim} \PY{o}{=} \PY{n}{ax1}\PY{o}{.}\PY{n}{get\PYZus{}xlim}\PY{p}{(}\PY{p}{)}
\PY{n}{x}\PY{o}{=}\PY{n}{y}\PY{o}{=}\PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{o}{*}\PY{n}{ylim}\PY{p}{)}
\PY{n}{ax1}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,}\PY{n}{y}\PY{p}{,}\PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}\PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax1}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{n}{ylim}\PY{p}{)}
\PY{n}{ax1}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{n}{xlim}\PY{p}{)}

\PY{n}{ax2}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}
    \PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}dr3}\PY{p}{,} 
    \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}delta\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{xerr} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}error\PYZus{}dr3}\PY{p}{,}
    \PY{n}{yerr} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}delta\PYZus{}err\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{none}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{color}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{k}\PY{l+s+s2}{\PYZdq{}}
\PY{p}{)}

\PY{n}{ax2}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta\PYZdl{} (GAIA DR2)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax2}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta\PYZdl{} (GAIA DR3)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{ylim} \PY{o}{=} \PY{n}{ax2}\PY{o}{.}\PY{n}{get\PYZus{}ylim}\PY{p}{(}\PY{p}{)}
\PY{n}{xlim} \PY{o}{=} \PY{n}{ax2}\PY{o}{.}\PY{n}{get\PYZus{}xlim}\PY{p}{(}\PY{p}{)}
\PY{n}{x}\PY{o}{=}\PY{n}{y}\PY{o}{=}\PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{o}{*}\PY{n}{ylim}\PY{p}{)}
\PY{n}{ax2}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,}\PY{n}{y}\PY{p}{,}\PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}\PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax2}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{n}{ylim}\PY{p}{)}
\PY{n}{ax2}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{n}{xlim}\PY{p}{)}

\PY{c+c1}{\PYZsh{} Then a histogram of the difference in the uncertainties}
\PY{n}{err\PYZus{}diff} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}error\PYZus{}dr3} \PY{o}{\PYZhy{}} \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}alpha\PYZus{}err\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}
\PY{n}{med} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmedian}\PY{p}{(}\PY{n}{err\PYZus{}diff}\PY{p}{)}
\PY{n}{ax3}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{err\PYZus{}diff}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fill}\PY{o}{=}\PY{k+kc}{False}\PY{p}{)}
\PY{n}{ax3}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{Delta }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha,DR3\PYZcb{} \PYZhy{} }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{Delta }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha,DR2\PYZcb{}\PYZdl{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax3}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{N}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax3}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{med}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax3}\PY{o}{.}\PY{n}{text}\PY{p}{(}\PY{n}{med}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{,} \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Median=}\PY{l+s+si}{\PYZob{}}\PY{n}{med}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{rotation}\PY{o}{=}\PY{l+m+mi}{90}\PY{p}{,} \PY{n}{verticalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{horizontalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{backgroundcolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{white}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{err\PYZus{}diff} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}error\PYZus{}dr3} \PY{o}{\PYZhy{}} \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}delta\PYZus{}err\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}
\PY{n}{med} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmedian}\PY{p}{(}\PY{n}{err\PYZus{}diff}\PY{p}{)}
\PY{n}{ax4}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{err\PYZus{}diff}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fill}\PY{o}{=}\PY{k+kc}{False}\PY{p}{)}
\PY{n}{ax4}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{Delta }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta,DR3\PYZcb{} \PYZhy{} }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{Delta }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta,DR2\PYZcb{}\PYZdl{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax4}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{N}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax4}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{med}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax4}\PY{o}{.}\PY{n}{text}\PY{p}{(}\PY{n}{med}\PY{p}{,} \PY{l+m+mf}{6.5}\PY{p}{,} \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Median=}\PY{l+s+si}{\PYZob{}}\PY{n}{med}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{rotation}\PY{o}{=}\PY{l+m+mi}{90}\PY{p}{,} \PY{n}{verticalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{horizontalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{backgroundcolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{white}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{c+c1}{\PYZsh{} then a histogram of the difference in the proper motion values}
\PY{n}{diff} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}dr3} \PY{o}{\PYZhy{}} \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}alpha\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}
\PY{n}{mean} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{diff}\PY{p}{)}
\PY{n}{std} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanstd}\PY{p}{(}\PY{n}{diff}\PY{p}{)}

\PY{n}{ax5}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{diff}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fill}\PY{o}{=}\PY{k+kc}{False}\PY{p}{)}
\PY{n}{ax5}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha,DR3\PYZcb{} \PYZhy{} }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha,DR2\PYZcb{}\PYZdl{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax5}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{N}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax5}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{mean}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax5}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{mean}\PY{o}{+}\PY{l+m+mi}{5}\PY{o}{*}\PY{n}{std}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{:}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax5}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{mean}\PY{o}{\PYZhy{}}\PY{l+m+mi}{5}\PY{o}{*}\PY{n}{std}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{:}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax5}\PY{o}{.}\PY{n}{text}\PY{p}{(}\PY{n}{mean}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{,} \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Mean=}\PY{l+s+si}{\PYZob{}}\PY{n}{mean}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{rotation}\PY{o}{=}\PY{l+m+mi}{90}\PY{p}{,} \PY{n}{verticalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{horizontalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{backgroundcolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{white}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax5}\PY{o}{.}\PY{n}{text}\PY{p}{(}\PY{n}{mean}\PY{o}{+}\PY{l+m+mi}{5}\PY{o}{*}\PY{n}{std}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{,} \PY{l+s+sa}{rf}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}5}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{sigma\PYZdl{}=}\PY{l+s+si}{\PYZob{}}\PY{n}{std}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{rotation}\PY{o}{=}\PY{l+m+mi}{90}\PY{p}{,} \PY{n}{verticalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{horizontalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{backgroundcolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{white}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{diff} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}dr3} \PY{o}{\PYZhy{}} \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}delta\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}
\PY{n}{mean} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{diff}\PY{p}{)}
\PY{n}{std} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanstd}\PY{p}{(}\PY{n}{diff}\PY{p}{)}

\PY{n}{ax6}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{diff}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fill}\PY{o}{=}\PY{k+kc}{False}\PY{p}{)}
\PY{n}{ax6}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta,DR3\PYZcb{} \PYZhy{} }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta,DR2\PYZcb{}\PYZdl{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax6}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{N}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax6}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{mean}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax6}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{mean}\PY{o}{+}\PY{l+m+mi}{5}\PY{o}{*}\PY{n}{std}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{:}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax6}\PY{o}{.}\PY{n}{axvline}\PY{p}{(}\PY{n}{mean}\PY{o}{\PYZhy{}}\PY{l+m+mi}{5}\PY{o}{*}\PY{n}{std}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{:}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\PY{n}{ax6}\PY{o}{.}\PY{n}{text}\PY{p}{(}\PY{n}{mean}\PY{p}{,} \PY{l+m+mf}{6.5}\PY{p}{,} \PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Mean=}\PY{l+s+si}{\PYZob{}}\PY{n}{mean}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{rotation}\PY{o}{=}\PY{l+m+mi}{90}\PY{p}{,} \PY{n}{verticalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{horizontalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{backgroundcolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{white}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax6}\PY{o}{.}\PY{n}{text}\PY{p}{(}\PY{n}{mean}\PY{o}{+}\PY{l+m+mi}{5}\PY{o}{*}\PY{n}{std}\PY{p}{,} \PY{l+m+mf}{6.5}\PY{p}{,} \PY{l+s+sa}{rf}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}5}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{sigma\PYZdl{}=}\PY{l+s+si}{\PYZob{}}\PY{n}{std}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{rotation}\PY{o}{=}\PY{l+m+mi}{90}\PY{p}{,} \PY{n}{verticalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{horizontalalignment}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{center}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{backgroundcolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{white}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{6}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
Text(10.219034500144446, 6.5, '\$5\textbackslash{}\textbackslash{}sigma\$=2.11')
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Franz_ASTR540_HW4_NB_files/Franz_ASTR540_HW4_NB_8_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsubsection{Problem 1b Discussion}\label{problem-1b-discussion}

The median improvement in the uncertainties in proper motions is
\textasciitilde0.50 mas/yr. See the above histograms (row 2 in the above
plot) for the distribution of improvements. As shown in the third row of
the above plot, there have not been any \textgreater{} 5 changes in the
proper motion between gaia DR3 and DR2 for this small sample.

\subsubsection{Problem 1c}\label{problem-1c}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{7}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} compute the error weighted mean}
\PY{n}{pmra}\PY{p}{,} \PY{n}{pmra\PYZus{}wt} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}dr3}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}dr3}\PY{p}{)}\PY{p}{]}\PY{p}{,} \PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}error\PYZus{}dr3}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}dr3}\PY{p}{)}\PY{p}{]}
\PY{n}{pmdec}\PY{p}{,} \PY{n}{pmdec\PYZus{}wt} \PY{o}{=} \PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}dr3}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}dr3}\PY{p}{)}\PY{p}{]}\PY{p}{,} \PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}error\PYZus{}dr3}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{pd}\PY{o}{.}\PY{n}{isna}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}dr3}\PY{p}{)}\PY{p}{]}

\PY{n}{where\PYZus{}not\PYZus{}outlier} \PY{o}{=} \PY{n}{pmra}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{cos}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{dec\PYZus{}dr3} \PY{o}{*} \PY{n}{np}\PY{o}{.}\PY{n}{pi}\PY{o}{/}\PY{l+m+mi}{180}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{o}{\PYZhy{}}\PY{l+m+mi}{4}
\PY{n}{pmra} \PY{o}{=} \PY{n}{pmra}\PY{p}{[}\PY{n}{where\PYZus{}not\PYZus{}outlier}\PY{p}{]}
\PY{n}{pmra\PYZus{}wt} \PY{o}{=} \PY{n}{pmra\PYZus{}wt}\PY{p}{[}\PY{n}{where\PYZus{}not\PYZus{}outlier}\PY{p}{]}
\PY{n}{pmdec} \PY{o}{=} \PY{n}{pmdec}\PY{p}{[}\PY{n}{where\PYZus{}not\PYZus{}outlier}\PY{p}{]}
\PY{n}{pmdec\PYZus{}wt} \PY{o}{=} \PY{n}{pmdec\PYZus{}wt}\PY{p}{[}\PY{n}{where\PYZus{}not\PYZus{}outlier}\PY{p}{]}

\PY{n}{mean\PYZus{}pmra} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{average}\PY{p}{(}\PY{n}{pmra}\PY{p}{,} \PY{n}{weights}\PY{o}{=}\PY{n}{pmra\PYZus{}wt}\PY{p}{)}
\PY{n}{std\PYZus{}pmra} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{average}\PY{p}{(}\PY{p}{(}\PY{n}{pmra}\PY{o}{\PYZhy{}}\PY{n}{mean\PYZus{}pmra}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{,} \PY{n}{weights}\PY{o}{=}\PY{n}{pmra\PYZus{}wt}\PY{p}{)}\PY{p}{)}

\PY{n}{mean\PYZus{}pmdec} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{average}\PY{p}{(}\PY{n}{pmdec}\PY{p}{,} \PY{n}{weights}\PY{o}{=}\PY{n}{pmdec\PYZus{}wt}\PY{p}{)}
\PY{n}{std\PYZus{}pmdec} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{average}\PY{p}{(}\PY{p}{(}\PY{n}{pmdec}\PY{o}{\PYZhy{}}\PY{n}{mean\PYZus{}pmdec}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{,} \PY{n}{weights}\PY{o}{=}\PY{n}{pmdec\PYZus{}wt}\PY{p}{)}\PY{p}{)}

\PY{n+nb}{print}\PY{p}{(}\PY{l+s+sa}{rf}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{(\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}a\PYZti{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{cos}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta\PYZdl{}, \PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta\PYZdl{}) = (}\PY{l+s+si}{\PYZob{}}\PY{n}{mean\PYZus{}pmra}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{, }\PY{l+s+si}{\PYZob{}}\PY{n}{mean\PYZus{}pmdec}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{) \PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{pm\PYZdl{} (}\PY{l+s+si}{\PYZob{}}\PY{n}{std\PYZus{}pmra}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{, }\PY{l+s+si}{\PYZob{}}\PY{n}{std\PYZus{}pmdec}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{c+c1}{\PYZsh{} make plot similar to Fig. 1 from the paper}

\PY{n}{fig}\PY{p}{,} \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplots}\PY{p}{(}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}
    \PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}dr3}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{cos}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{dec\PYZus{}dr3} \PY{o}{*} \PY{n}{np}\PY{o}{.}\PY{n}{pi}\PY{o}{/}\PY{l+m+mi}{180}\PY{p}{)}\PY{p}{,}
    \PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}dr3}\PY{p}{,}
    \PY{n}{xerr}\PY{o}{=}\PY{n}{data}\PY{o}{.}\PY{n}{pmra\PYZus{}error\PYZus{}dr3}\PY{p}{,}
    \PY{n}{yerr}\PY{o}{=}\PY{n}{data}\PY{o}{.}\PY{n}{pmdec\PYZus{}error\PYZus{}dr3}\PY{p}{,}
    \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{none}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{marker}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{o}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} 
    \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{GAIA DR3}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{cornflowerblue}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}
    \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.5}
\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}
    \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}alpha\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{cos}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{delta\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)} \PY{o}{*} \PY{n}{np}\PY{o}{.}\PY{n}{pi}\PY{o}{/}\PY{l+m+mi}{180}\PY{p}{)}\PY{p}{,}
    \PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}delta\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{xerr}\PY{o}{=}\PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}alpha\PYZus{}err\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{yerr}\PY{o}{=}\PY{n}{data}\PY{o}{.}\PY{n}{mu\PYZus{}delta\PYZus{}err\PYZus{}dr2}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{none}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{marker}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{o}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}
    \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{GAIA DR2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
    \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{firebrick}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}
    \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.5}
\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta\PYZdl{} [mas/yr]}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{mu\PYZus{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha\PYZti{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{cos}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{delta\PYZdl{} [mas/yr]}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{8}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{8}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
(\$\textbackslash{}mu\_a\textasciitilde{}\textbackslash{}cos\textbackslash{}delta\$, \$\textbackslash{}mu\_\textbackslash{}delta\$) = (-1.39, -1.44) \$\textbackslash{}pm\$ (0.69, 0.92)
    \end{Verbatim}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{7}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
<matplotlib.legend.Legend at 0x7d2d117a41d0>
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Franz_ASTR540_HW4_NB_files/Franz_ASTR540_HW4_NB_10_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsubsection{Problem 1c Discussion}\label{problem-1c-discussion}

I get a mean proper motion, weighted by the uncertainties, of
(\(\mu_a~\cos\delta\), \(\mu_\delta\)) = (-1.39, -1.44) \(\pm\) (0.69,
0.92) from the Gaia DR3 dataset. This is consistent with the value from
Gaia DR2 within \(1\sigma\). Although, interestingly, the uncertainties
on the mean are much larger for DR3 than DR2. I'm not sure how it was
computed in the paper, but I did did a weighted standard deviation to
compute the statistical uncertainty.

\subsubsection{Problem 1d}\label{problem-1d}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{8}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{isofile} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fehm20afem2.SDSSugriz}\PY{l+s+s2}{\PYZdq{}}
\PY{n}{boo3\PYZus{}distance} \PY{o}{=} \PY{l+m+mf}{46.5} \PY{c+c1}{\PYZsh{} kpc, from the paper}

\PY{k}{def} \PY{n+nf}{M\PYZus{}to\PYZus{}m}\PY{p}{(}\PY{n}{M}\PY{p}{,} \PY{n}{d}\PY{o}{=}\PY{n}{boo3\PYZus{}distance}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{n}{M} \PY{o}{+} \PY{l+m+mi}{5}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{log10}\PY{p}{(}\PY{n}{d}\PY{o}{/}\PY{l+m+mi}{10}\PY{p}{)}

\PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{n}{isofile}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{k}{as} \PY{n}{f}\PY{p}{:}
    \PY{n}{iso\PYZus{}data} \PY{o}{=} \PY{n}{f}\PY{o}{.}\PY{n}{readlines}\PY{p}{(}\PY{p}{)}

\PY{n}{hdr} \PY{o}{=} \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{:}\PY{l+m+mi}{6}\PY{p}{]}
\PY{n}{iso\PYZus{}data} \PY{o}{=} \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{l+m+mi}{6}\PY{p}{:}\PY{p}{]}

\PY{n}{hdr\PYZus{}vals} \PY{o}{=} \PY{p}{[}\PY{n}{val}\PY{o}{.}\PY{n}{strip}\PY{p}{(}\PY{p}{)} \PY{k}{for} \PY{n}{val} \PY{o+ow}{in} \PY{n}{hdr}\PY{p}{[}\PY{l+m+mi}{3}\PY{p}{]}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ }\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{k}{if} \PY{n}{val} \PY{o+ow}{not} \PY{o+ow}{in} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{\PYZcb{}}\PY{p}{]}
\PY{n}{mix\PYZus{}len}\PY{p}{,} \PY{n}{Y}\PY{p}{,} \PY{n}{Z}\PY{p}{,} \PY{n}{Zeff}\PY{p}{,} \PY{n}{Fe\PYZus{}H}\PY{p}{,} \PY{n}{a\PYZus{}Fe} \PY{o}{=} \PY{n}{hdr\PYZus{}vals}

\PY{n}{isochrones\PYZus{}in\PYZus{}this\PYZus{}file} \PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{n}{base\PYZus{}iso} \PY{o}{=} \PY{n+nb}{dict}\PY{p}{(}
    \PY{n}{idx} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{M} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{LogTeff} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{LogG} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{LogL} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{u} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{g} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{r} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{i} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{z} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{age} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{mix\PYZus{}len} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{Y} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{Z} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{Zeff} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{Fe\PYZus{}H} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
    \PY{n}{a\PYZus{}Fe} \PY{o}{=} \PY{p}{[}\PY{p}{]}\PY{p}{,}
\PY{p}{)}
\PY{n}{iso} \PY{o}{=} \PY{n}{deepcopy}\PY{p}{(}\PY{n}{base\PYZus{}iso}\PY{p}{)}
\PY{n}{age} \PY{o}{=} \PY{k+kc}{None}

\PY{k}{for} \PY{n}{j} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{iso\PYZus{}data}\PY{p}{)}\PY{p}{)}\PY{p}{:}
    \PY{k}{if} \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{j}\PY{p}{]} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
        \PY{k}{continue}

    \PY{k}{if} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}AGE}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{in} \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{j}\PY{p}{]}\PY{p}{:}
        \PY{n}{isochrones\PYZus{}in\PYZus{}this\PYZus{}file}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{iso}\PY{p}{)}\PY{p}{)}

        \PY{n}{age} \PY{o}{=} \PY{n+nb}{float}\PY{p}{(}\PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{j}\PY{p}{]}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ EEPS=}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{AGE=}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{strip}\PY{p}{(}\PY{p}{)}\PY{p}{)}
        \PY{n}{iso} \PY{o}{=} \PY{n}{deepcopy}\PY{p}{(}\PY{n}{base\PYZus{}iso}\PY{p}{)}
        \PY{k}{continue}

    \PY{k}{if} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}}\PY{l+s+s1}{\PYZsq{}} \PY{o}{==} \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{j}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{:}
        \PY{c+c1}{\PYZsh{} this line is a header}
        \PY{k}{continue}

    \PY{n}{line} \PY{o}{=} \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{j}\PY{p}{]}
    \PY{n}{goodline} \PY{o}{=} \PY{p}{[}\PY{n}{val} \PY{k}{for} \PY{n}{val} \PY{o+ow}{in} \PY{n}{line}\PY{o}{.}\PY{n}{strip}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ }\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{k}{if} \PY{n+nb}{len}\PY{p}{(}\PY{n}{val}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{]}

    \PY{k}{if} \PY{n+nb}{len}\PY{p}{(}\PY{n}{goodline}\PY{p}{)} \PY{o}{!=} \PY{l+m+mi}{10}\PY{p}{:}
        \PY{n+nb}{print}\PY{p}{(}\PY{n}{goodline}\PY{p}{)}
        \PY{k}{continue}

    \PY{k}{for} \PY{n}{val}\PY{p}{,} \PY{n}{key} \PY{o+ow}{in} \PY{n+nb}{zip}\PY{p}{(}\PY{n}{goodline}\PY{p}{,} \PY{n}{iso}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{:}
        \PY{c+c1}{\PYZsh{} if key == \PYZsq{}age\PYZsq{}: continue}
        \PY{n}{iso}\PY{p}{[}\PY{n}{key}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{val}\PY{p}{)}

    \PY{n}{iso}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{age}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{age}\PY{p}{)}
    \PY{n}{iso}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mix\PYZus{}len}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{mix\PYZus{}len}\PY{p}{)}
    \PY{n}{iso}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Y}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{Y}\PY{p}{)}
    \PY{n}{iso}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Z}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{Z}\PY{p}{)}
    \PY{n}{iso}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Zeff}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{Zeff}\PY{p}{)}
    \PY{n}{iso}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Fe\PYZus{}H}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{Fe\PYZus{}H}\PY{p}{)}
    \PY{n}{iso}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{a\PYZus{}Fe}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{a\PYZus{}Fe}\PY{p}{)}

\PY{n}{iso\PYZus{}data} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{concat}\PY{p}{(}\PY{n}{isochrones\PYZus{}in\PYZus{}this\PYZus{}file}\PY{p}{)}\PY{o}{.}\PY{n}{reset\PYZus{}index}\PY{p}{(}\PY{n}{drop}\PY{o}{=}\PY{k+kc}{True}\PY{p}{)}

\PY{c+c1}{\PYZsh{} select just the 10 Gyr one}
\PY{n}{iso\PYZus{}data} \PY{o}{=} \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{age} \PY{o}{==} \PY{l+m+mi}{10}\PY{p}{]}

\PY{c+c1}{\PYZsh{} scale all of the magnitudes to boo3\PYZus{}distance}
\PY{k}{for} \PY{n}{filt} \PY{o+ow}{in} \PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{u}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{g}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{i}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{z}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{:}
    \PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{filt}\PY{p}{]} \PY{o}{=} \PY{n}{M\PYZus{}to\PYZus{}m}\PY{p}{(}\PY{n}{iso\PYZus{}data}\PY{p}{[}\PY{n}{filt}\PY{p}{]}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,} \PY{n}{d}\PY{o}{=}\PY{n}{boo3\PYZus{}distance}\PY{o}{*}\PY{l+m+mf}{1e3}\PY{p}{)}

\PY{n}{iso\PYZus{}data}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{8}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
      idx         M LogTeff    LogG     LogL          u          g          r  \textbackslash{}
7068   12  0.103511  3.5376  5.3748  -2.8181  36.404765  32.468665  30.781265
7069   13  0.113801  3.5502  5.3310  -2.6829  35.490765  31.976365  30.362465
7070   14  0.125317  3.5628  5.2860  -2.5455  34.679065  31.476965  29.948765
7071   15  0.138532  3.5743  5.2421  -2.4122  33.980765  30.999365  29.558465
7072   16  0.152713  3.5827  5.2041  -2.2982  33.447165  30.609265  29.235965
{\ldots}   {\ldots}       {\ldots}     {\ldots}     {\ldots}      {\ldots}        {\ldots}        {\ldots}        {\ldots}
7331  275  0.856466  3.6502  0.7935   3.1312  18.666965  16.462765  15.466865
7332  276  0.856473  3.6486  0.7629   3.1556  18.667265  16.425365  15.414365
7333  277  0.856479  3.6471  0.7331   3.1795  18.669365  16.389665  15.363365
7334  278  0.856486  3.6457  0.7042   3.2026  18.672265  16.355565  15.314365
7335  279  0.856491  3.6444  0.6775   3.2240  18.675365  16.324365  15.269265

              i          z   age mix\_len       Y           Z        Zeff  \textbackslash{}
7068  30.051865  29.662965  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7069  29.705465  29.342965  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7070  29.348765  29.010965  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7071  29.001365  28.684265  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7072  28.706265  28.403265  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
{\ldots}         {\ldots}        {\ldots}   {\ldots}     {\ldots}     {\ldots}         {\ldots}         {\ldots}
7331  15.087665  14.876465  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7332  15.029165  14.814465  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7333  14.972265  14.754065  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7334  14.917465  14.695765  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04
7335  14.866865  14.641965  10.0  1.9380  0.2452  1.2990E-04  1.6990E-04

       Fe\_H   a\_Fe
7068  -2.01  -0.20
7069  -2.01  -0.20
7070  -2.01  -0.20
7071  -2.01  -0.20
7072  -2.01  -0.20
{\ldots}     {\ldots}    {\ldots}
7331  -2.01  -0.20
7332  -2.01  -0.20
7333  -2.01  -0.20
7334  -2.01  -0.20
7335  -2.01  -0.20

[268 rows x 17 columns]
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{9}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{boo3\PYZus{}distance} \PY{o}{=} \PY{l+m+mf}{46.5} \PY{c+c1}{\PYZsh{} kpc, from the paper}
\PY{n}{boo3\PYZus{}delta\PYZus{}distance} \PY{o}{=} \PY{l+m+mi}{2} \PY{c+c1}{\PYZsh{} kpc, the value used in the paper}
\PY{n}{boo3\PYZus{}radius} \PY{o}{=} \PY{l+m+mi}{30} \PY{c+c1}{\PYZsh{} arcmin, value used for the cone search in the paper}
\PY{n}{pmra\PYZus{}min}\PY{p}{,} \PY{n}{pmra\PYZus{}max} \PY{o}{=} \PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{3}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)} \PY{c+c1}{\PYZsh{} mu\PYZus{}alpha * cos(delta) in mas/yr}
\PY{n}{pmdec\PYZus{}min}\PY{p}{,} \PY{n}{pmdec\PYZus{}max} \PY{o}{=} \PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{3}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)} \PY{c+c1}{\PYZsh{} mu\PYZus{}delta in mas/yr}

\PY{c+c1}{\PYZsh{} Color\PYZhy{}magnitude filter: this requires stars to be within}
\PY{c+c1}{\PYZsh{} 0.1 mag at g0 = 16, increasing linearly to 0.2 mag width at g0 = 22.5}
\PY{k}{def} \PY{n+nf}{cmd\PYZus{}filter}\PY{p}{(}\PY{n}{g\PYZus{}mag}\PY{p}{)}\PY{p}{:}
    \PY{n}{p} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{polyfit}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{l+m+mi}{16}\PY{p}{,} \PY{l+m+mf}{22.5}\PY{p}{)}\PY{p}{,} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{l+m+mf}{0.1}\PY{p}{,}\PY{l+m+mf}{0.2}\PY{p}{)}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
    \PY{n}{cmd\PYZus{}filter} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{polyval}\PY{p}{(}\PY{n}{p}\PY{o}{=}\PY{n}{p}\PY{p}{,} \PY{n}{x}\PY{o}{=}\PY{n}{g\PYZus{}mag}\PY{p}{)}
    \PY{k}{return} \PY{n}{cmd\PYZus{}filter}

\PY{k}{def} \PY{n+nf}{is\PYZus{}within\PYZus{}cmd\PYZus{}filter}\PY{p}{(}\PY{n}{g\PYZus{}mag}\PY{p}{,} \PY{n}{r\PYZus{}mag}\PY{p}{,} \PY{n}{isochrone}\PY{p}{)}\PY{p}{:}
    \PY{n}{filt} \PY{o}{=} \PY{n}{cmd\PYZus{}filter}\PY{p}{(}\PY{n}{g\PYZus{}mag}\PY{p}{)}

    \PY{c+c1}{\PYZsh{} find closest isochrone point}
    \PY{c+c1}{\PYZsh{} good enough for now}
    \PY{n}{min\PYZus{}idx} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{argmin}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{g\PYZus{}mag}\PY{o}{\PYZhy{}}\PY{n}{isochrone}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{values}\PY{p}{)}\PY{p}{)}
    \PY{n}{g}\PY{p}{,} \PY{n}{r} \PY{o}{=} \PY{n}{isochrone}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{values}\PY{p}{[}\PY{n}{min\PYZus{}idx}\PY{p}{]}\PY{p}{,} \PY{n}{isochrone}\PY{o}{.}\PY{n}{r}\PY{o}{.}\PY{n}{values}\PY{p}{[}\PY{n}{min\PYZus{}idx}\PY{p}{]}

    \PY{n}{color} \PY{o}{=} \PY{n}{g\PYZus{}mag} \PY{o}{\PYZhy{}} \PY{n}{r\PYZus{}mag}

    \PY{k}{return} \PY{n}{color} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{p}{(}\PY{n}{g}\PY{o}{\PYZhy{}}\PY{n}{r}\PY{o}{+}\PY{n}{filt}\PY{p}{)} \PY{o+ow}{and} \PY{n}{color} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{p}{(}\PY{n}{g}\PY{o}{\PYZhy{}}\PY{n}{r}\PY{o}{\PYZhy{}}\PY{n}{filt}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{10}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} first do a cone search in Gaia DR3}
\PY{n}{Gaia}\PY{o}{.}\PY{n}{ROW\PYZus{}LIMIT} \PY{o}{=} \PY{l+m+mi}{100\PYZus{}000}
\PY{n}{boo3\PYZus{}coord} \PY{o}{=} \PY{n}{SkyCoord}\PY{p}{(}\PY{l+m+mf}{209.3}\PY{p}{,} \PY{l+m+mf}{26.8}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{n}{u}\PY{o}{.}\PY{n}{deg}\PY{p}{)}
\PY{n}{res} \PY{o}{=} \PY{n}{Gaia}\PY{o}{.}\PY{n}{query\PYZus{}object}\PY{p}{(}
    \PY{n}{coordinate}\PY{o}{=}\PY{n}{boo3\PYZus{}coord}\PY{p}{,} 
    \PY{n}{radius}\PY{o}{=}\PY{n}{boo3\PYZus{}radius}\PY{o}{*}\PY{n}{u}\PY{o}{.}\PY{n}{arcmin}\PY{p}{,}
    \PY{n}{verbose}\PY{o}{=}\PY{k+kc}{False}
\PY{p}{)}

\PY{c+c1}{\PYZsh{} filter on proper motion}
\PY{n}{ra\PYZus{}res} \PY{o}{=} \PY{n}{res}\PY{p}{[}\PY{p}{(}\PY{n}{res}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pmra}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{cos}\PY{p}{(}\PY{n}{res}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ra}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{to}\PY{p}{(}\PY{n}{u}\PY{o}{.}\PY{n}{radian}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{n}{pmra\PYZus{}min}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n}{res}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pmra}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{cos}\PY{p}{(}\PY{n}{res}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ra}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{to}\PY{p}{(}\PY{n}{u}\PY{o}{.}\PY{n}{radian}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZlt{}} \PY{n}{pmra\PYZus{}max}\PY{p}{)}\PY{p}{]}
\PY{n}{dec\PYZus{}res} \PY{o}{=} \PY{n}{ra\PYZus{}res}\PY{p}{[}\PY{p}{(}\PY{n}{ra\PYZus{}res}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pmdec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{\PYZgt{}} \PY{n}{pmdec\PYZus{}min}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n}{ra\PYZus{}res}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pmdec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{\PYZlt{}} \PY{n}{pmdec\PYZus{}max}\PY{p}{)}\PY{p}{]}

\PY{c+c1}{\PYZsh{} filter on the CMD cuts}
\PY{n}{res} \PY{o}{=} \PY{n}{get\PYZus{}sdss\PYZus{}colors}\PY{p}{(}\PY{n}{dec\PYZus{}res}\PY{o}{.}\PY{n}{to\PYZus{}pandas}\PY{p}{(}\PY{p}{)}\PY{p}{)}

\PY{n}{res}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{is\PYZus{}within\PYZus{}cmd\PYZus{}filter}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{n}{res}\PY{o}{.}\PY{n}{apply}\PY{p}{(}
    \PY{k}{lambda} \PY{n}{row} \PY{p}{:} \PY{n}{is\PYZus{}within\PYZus{}cmd\PYZus{}filter}\PY{p}{(}\PY{n}{row}\PY{o}{.}\PY{n}{g\PYZus{}sdss}\PY{p}{,} \PY{n}{row}\PY{o}{.}\PY{n}{r\PYZus{}sdss}\PY{p}{,} \PY{n}{iso\PYZus{}data}\PY{p}{)}\PY{p}{,}
    \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}
\PY{p}{)}

\PY{n}{res} \PY{o}{=} \PY{n}{res}\PY{p}{[}\PY{n}{res}\PY{o}{.}\PY{n}{is\PYZus{}within\PYZus{}cmd\PYZus{}filter}\PY{p}{]}

\PY{c+c1}{\PYZsh{} filter on g\PYZus{}0}
\PY{n}{res} \PY{o}{=} \PY{n}{res}\PY{p}{[}\PY{n}{res}\PY{o}{.}\PY{n}{g\PYZus{}sdss} \PY{o}{\PYZlt{}} \PY{l+m+mi}{20}\PY{p}{]}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
/tmp/ipykernel\_466903/2148086655.py:22: UserWarning: Skipping Gaia DR3
1450837238169921280 because no SDSS data!
  warnings.warn(f"Skipping \{row.DESIGNATION\} because no SDSS data!")
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{11}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} read in Table 2 from Carlin \PYZam{} Sand (2018)}
\PY{n}{dr2\PYZus{}res} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{read\PYZus{}csv}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Carlin\PYZus{}and\PYZus{}Sand\PYZus{}2018\PYZus{}Table2.txt}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{sep}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}t}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{dr2\PYZus{}res} \PY{o}{=} \PY{n}{get\PYZus{}sdss\PYZus{}colors}\PY{p}{(}\PY{n}{dr2\PYZus{}res}\PY{p}{,} \PY{n}{ra\PYZus{}key}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{alpha}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{dec\PYZus{}key}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{delta}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{c+c1}{\PYZsh{} then plot it up!}
\PY{n}{fig}\PY{p}{,} \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplots}\PY{p}{(}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{r}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,} \PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{10 Gyr}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{[Fe/H]=\PYZhy{}2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{g\PYZus{}sdss}\PY{o}{\PYZhy{}}\PY{n}{data}\PY{o}{.}\PY{n}{r\PYZus{}sdss}\PY{p}{,} \PY{n}{data}\PY{o}{.}\PY{n}{g\PYZus{}sdss}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{o}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{none}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Original Data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{res}\PY{o}{.}\PY{n}{g\PYZus{}sdss}\PY{o}{\PYZhy{}}\PY{n}{res}\PY{o}{.}\PY{n}{r\PYZus{}sdss}\PY{p}{,} \PY{n}{res}\PY{o}{.}\PY{n}{g\PYZus{}sdss}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{s}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{edgecolor}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{facecolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{none}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Gaia DR3 Search Results}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{dr2\PYZus{}res}\PY{o}{.}\PY{n}{g\PYZus{}sdss}\PY{o}{\PYZhy{}}\PY{n}{dr2\PYZus{}res}\PY{o}{.}\PY{n}{r\PYZus{}sdss}\PY{p}{,} \PY{n}{dr2\PYZus{}res}\PY{o}{.}\PY{n}{g\PYZus{}sdss}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{X}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{edgecolor}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{facecolor}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{none}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Gaia DR2 Search Results}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Carlin \PYZam{} Sand (2018)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}
    \PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{r}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)} \PY{o}{+} \PY{n}{cmd\PYZus{}filter}\PY{p}{(}\PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{)}\PY{p}{,}
    \PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}
    \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}
    \PY{n}{label} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{CMD Filter Ranges}\PY{l+s+s2}{\PYZdq{}}
\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}
    \PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{r}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{cmd\PYZus{}filter}\PY{p}{(}\PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{)}\PY{p}{,}
    \PY{n}{iso\PYZus{}data}\PY{o}{.}\PY{n}{g}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{n+nb}{float}\PY{p}{)}\PY{p}{,}
    \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}
    \PY{n}{color}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}
\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{invert\PYZus{}yaxis}\PY{p}{(}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{g\PYZhy{}r}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{g}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{10}\PY{p}{)}

\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{l+m+mi}{22}\PY{p}{,} \PY{l+m+mi}{16}\PY{p}{)}
\PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mf}{0.8}\PY{p}{,} \PY{l+m+mf}{1.8}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{11}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
(-0.8, 1.8)
\end{Verbatim}
\end{tcolorbox}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Franz_ASTR540_HW4_NB_files/Franz_ASTR540_HW4_NB_15_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsubsection{Problem 1d Discussion}\label{problem-1d-discussion}

We used the same proper motion filter, same color-magnitude diagram
filter, and maximum \(g < 20\) from Carlin \& Sand (2018) to search for
plausible main sequence and red giant branch proper motion members of
Bootes III. The results are shown in the figure above.

From the SDSS colors and Isochrones (from HW 1), we find 7 new proper
motion members of Bootes III. The Isochrone shown is for a 10 Gyr
cluster with a metallicity \([Fe/H]=-2\). Note that we do not find some
of the original members from Carlin \& Sand (2018) for two primary
reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We use SDSS magnitudes instead of Panstarrs magnitudes. This is
  because the Isochrone we have access to is for SDSS magnitudes. This
  changes the exact width of the Color Magnitude Diagram filter and
  eliminates some of them.
\item
  The Isochrone we are using does not include the population of bluer
  stars on the right side of the color-magnitude diagram. This is a
  limitation of this analysis.
\end{enumerate}

Additionally, it is important to note that these are simply
\textbf{plausible} members of Bootes III that need true RV measurements
to confirm. As described in Carlin \& Sand (2018), there are numerous
sources of contaminants for these proper motion calculations.

    \subsection{Problem 2}\label{problem-2}

Congratulations! You have discovered a dwarf galaxy around the Milky
Way. Wishing to refine it's distance, you took time series imaging data
of this new system and identified 3 RR Lyrae stars, whose proper- ties
are in the data file ``rrl.dat''. Infer the distance to the new dwarf
galaxy using the basic relations in Caceres \& Catelan 2008, ApJSS, 179,
242. You can assume a metallicity of Z=0.0001, but assume a 10\%
uncertainty on this value. For the average i and z band magnitudes, you
can assume 0.03 mag uncertainties, and a 0.01 day uncertainty on the RRL
periods. What is the distance (with uncertainty) to your newly
discovered dwarf galaxy?

The relevant equations in Caceres \& CAtelan (2008) are equations (6)
and (7):

\(M_z = 0.839 - 1.295~\log P + 0.211~\log Z\)

\(M_i = 0.908 - 1.035~\log P + 0.220~\log Z\)

Which has the general form
\(M_f = \gamma - \alpha~\log P + \beta~\log Z\)

From these we can use the distance modulus equation to find the distance
to each RR Lyrae Star

\(m - M = 5\log d - 5 \rightarrow \log d = \frac{m - M + 5}{5}\)

Substituting in the general formulation of M

\(\log d_f = \frac{1}{5} \left[ m_f - \gamma + \alpha~\log P - \beta~\log Z + 5 \right]\)

Which then, propagating uncertainty, has an uncertainty of

\(\sigma_d = d_f \left[\left(\alpha\frac{\sigma_P}{\ln(10)~ P}\right)^2 + \left(\beta\frac{\sigma_Z}{\ln(10)~Z}\right)^2 + \left(\frac{\sigma_m}{m_f}\right)^2 \right]^{1/2}\)

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{12}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{log\PYZus{}distance}\PY{p}{(}\PY{n}{P}\PY{p}{,} \PY{n}{sigma\PYZus{}P}\PY{p}{,} \PY{n}{Z}\PY{p}{,} \PY{n}{sigma\PYZus{}Z}\PY{p}{,} \PY{n}{m}\PY{p}{,} \PY{n}{sigma\PYZus{}m}\PY{p}{,} \PY{n}{alpha}\PY{p}{,} \PY{n}{beta}\PY{p}{,} \PY{n}{gamma}\PY{p}{)}\PY{p}{:}
    \PY{n}{log\PYZus{}d} \PY{o}{=} \PY{l+m+mi}{1}\PY{o}{/}\PY{l+m+mi}{5} \PY{o}{*} \PY{p}{(}\PY{n}{m} \PY{o}{\PYZhy{}} \PY{n}{gamma} \PY{o}{+} \PY{n}{alpha}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{log10}\PY{p}{(}\PY{n}{P}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{beta}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{log10}\PY{p}{(}\PY{n}{Z}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{5}\PY{p}{)}
    \PY{n}{sigma\PYZus{}d} \PY{o}{=} \PY{n}{log\PYZus{}d} \PY{o}{*} \PY{p}{(} \PY{p}{(}\PY{n}{alpha}\PY{o}{*}\PY{n}{sigma\PYZus{}P}\PY{o}{/}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{log}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{)}\PY{o}{*}\PY{n}{P}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2} \PY{o}{+} \PY{p}{(}\PY{n}{beta}\PY{o}{*}\PY{n}{sigma\PYZus{}Z}\PY{o}{/}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{log}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{)}\PY{o}{*}\PY{n}{Z}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2} \PY{o}{+} \PY{p}{(}\PY{n}{sigma\PYZus{}m}\PY{o}{/}\PY{n}{m}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2} \PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{l+m+mf}{0.5}

    \PY{k}{return} \PY{n}{log\PYZus{}d}\PY{p}{,} \PY{n}{sigma\PYZus{}d}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{13}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{rr\PYZus{}lyrae} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{read\PYZus{}csv}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{rrl.dat}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{sep}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ }\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}

\PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Z}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{l+m+mf}{0.0001}
\PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Z\PYZus{}err}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{l+m+mf}{0.1}\PY{o}{*}\PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{Z}

\PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{i\PYZus{}mag\PYZus{}err}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{l+m+mf}{0.03}
\PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{z\PYZus{}mag\PYZus{}err}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{l+m+mf}{0.03}

\PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Period\PYZus{}err}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{l+m+mf}{0.01}

\PY{n}{i\PYZus{}band\PYZus{}kwargs} \PY{o}{=} \PY{n+nb}{dict}\PY{p}{(}
    \PY{n}{alpha} \PY{o}{=} \PY{l+m+mf}{1.035}\PY{p}{,}
    \PY{n}{beta} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{l+m+mf}{0.220}\PY{p}{,}
    \PY{n}{gamma} \PY{o}{=} \PY{l+m+mf}{0.908}
\PY{p}{)}

\PY{n}{z\PYZus{}band\PYZus{}kwargs} \PY{o}{=} \PY{n+nb}{dict}\PY{p}{(}
    \PY{n}{alpha} \PY{o}{=} \PY{l+m+mf}{1.295}\PY{p}{,}
    \PY{n}{beta} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{l+m+mf}{0.211}\PY{p}{,}
    \PY{n}{gamma} \PY{o}{=} \PY{l+m+mf}{0.839}
\PY{p}{)}

\PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{log\PYZus{}d\PYZus{}i}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{log\PYZus{}d\PYZus{}i\PYZus{}err}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{log\PYZus{}distance}\PY{p}{(}
    \PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Period(d)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{Period\PYZus{}err}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{Z}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{Z\PYZus{}err}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{i\PYZus{}mag}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{i\PYZus{}mag\PYZus{}err}\PY{p}{,}
    \PY{o}{*}\PY{o}{*}\PY{n}{i\PYZus{}band\PYZus{}kwargs}
\PY{p}{)}

\PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{log\PYZus{}d\PYZus{}z}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{log\PYZus{}d\PYZus{}z\PYZus{}err}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{log\PYZus{}distance}\PY{p}{(}
    \PY{n}{rr\PYZus{}lyrae}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Period(d)}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{Period\PYZus{}err}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{Z}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{Z\PYZus{}err}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{z\PYZus{}mag}\PY{p}{,}
    \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{z\PYZus{}mag\PYZus{}err}\PY{p}{,}
    \PY{o}{*}\PY{o}{*}\PY{n}{z\PYZus{}band\PYZus{}kwargs}
\PY{p}{)}

\PY{n}{rr\PYZus{}lyrae}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{13}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
   Name     RA(deg)   DEC(deg)  Period(d)  i\_mag  z\_mag    A\_i    A\_z       Z  \textbackslash{}
0  RRL1  189.570323 -40.939879   0.389918  20.92  20.86  0.192  0.146  0.0001
1  RRL2  189.633635 -40.878072   0.422481  20.88  20.83  0.203  0.155  0.0001
2  RRL3  189.584351 -41.101214   0.735898  20.67  20.59  0.177  0.135  0.0001

     Z\_err  i\_mag\_err  z\_mag\_err  Period\_err   log\_d\_i  log\_d\_i\_err   log\_d\_z  \textbackslash{}
0  0.00001       0.03       0.03        0.01  4.741731     0.071321  4.729462
1  0.00001       0.03       0.03        0.01  4.740942     0.068136  4.732484
2  0.00001       0.03       0.03        0.01  4.748831     0.054291  4.746906

   log\_d\_z\_err
0     0.081105
1     0.076786
2     0.057062
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{14}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{print\PYZus{}results}\PY{p}{(}\PY{n}{log\PYZus{}dist}\PY{p}{,} \PY{n}{log\PYZus{}dist\PYZus{}err}\PY{p}{,} \PY{n}{varname}\PY{p}{)}\PY{p}{:}

    \PY{n}{dist} \PY{o}{=} \PY{l+m+mi}{10}\PY{o}{*}\PY{o}{*}\PY{n}{log\PYZus{}dist}
    \PY{n}{dist\PYZus{}err} \PY{o}{=} \PY{n}{log\PYZus{}dist\PYZus{}err} \PY{o}{*} \PY{n}{dist}\PY{o}{/}\PY{n}{log\PYZus{}dist}

    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+sa}{fr}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdl{}}\PY{l+s+si}{\PYZob{}}\PY{n}{varname}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{\PYZdl{} = }\PY{l+s+si}{\PYZob{}}\PY{n}{dist}\PY{o}{/}\PY{l+m+mf}{1e3}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ \PYZdl{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{pm\PYZdl{} }\PY{l+s+si}{\PYZob{}}\PY{n}{dist\PYZus{}err}\PY{o}{/}\PY{l+m+mf}{1e3}\PY{l+s+si}{:}\PY{l+s+s2}{.2f}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ kpc}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{n}{print\PYZus{}results}\PY{p}{(}\PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{log\PYZus{}d\PYZus{}z}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{log\PYZus{}d\PYZus{}z\PYZus{}err}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{d\PYZus{}z}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\PY{n}{print\PYZus{}results}\PY{p}{(}\PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{log\PYZus{}d\PYZus{}i}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{rr\PYZus{}lyrae}\PY{o}{.}\PY{n}{log\PYZus{}d\PYZus{}i\PYZus{}err}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{d\PYZus{}i}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\$d\_z\$ = 54.49 \$\textbackslash{}pm\$ 0.82 kpc
\$d\_i\$ = 55.44 \$\textbackslash{}pm\$ 0.75 kpc
    \end{Verbatim}

    \subsubsection{Problem 2 Discussion}\label{problem-2-discussion}

The full results of the distance calculations for these three RR Lyrae
stars are shown in the table above. Since I compute the distance in both
z-band and i-band independently, we get two measurements of the
distance. These measurements, which I get by averaging the log distance
results, are consistent within uncertainties and give a distance of
\(d\sim (55\pm1) kpc\).

    \subsection{Problem 3}\label{problem-3}

Black Hole Accretion Time Scales As a mass m of gas falls into a black
hole, at most 0.1mc2 is likely to emerge as radiation; the rest is
swallowed by the black hole. Show that the Eddington luminosity for a
black hole of mass M is equivalent to 2×10−9 M c2 yr−1. Explain why we
expect the black hole's mass to grow by at least a factor of e every 5 ×
107 years.

We start by deriving the eddington luminosity in a typical way

\begin{equation}
F_g = F_{rad}
\end{equation}

\begin{equation}
\frac{G M m_p}{r^2} = \frac{ \sigma_T L_{edd} }{ 4 \pi c r^2 }
\end{equation}

Solving for \(L_{edd}\)

\begin{equation}
L_{edd} = \frac{4\pi G m_p}{c \sigma_T} Mc^2$
\end{equation}

Or in terms of years

\begin{equation}
L_{edd} = \frac{1.261\times10^8 \pi G m_p}{c \sigma_T} \frac{M c^2}{yr}
\end{equation}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{17}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{sec\PYZus{}to\PYZus{}year} \PY{o}{=} \PY{l+m+mi}{3600} \PY{o}{*} \PY{l+m+mi}{24} \PY{o}{*} \PY{l+m+mi}{365} \PY{o}{*} \PY{n}{u}\PY{o}{.}\PY{n}{s} \PY{o}{/} \PY{n}{u}\PY{o}{.}\PY{n}{yr}
\PY{c+c1}{\PYZsh{} print(f\PYZdq{}\PYZob{}sec\PYZus{}to\PYZus{}year*4:.2e\PYZcb{}\PYZdq{})}

\PY{n}{res\PYZus{}factor} \PY{o}{=} \PY{l+m+mi}{4}\PY{o}{*}\PY{n}{sec\PYZus{}to\PYZus{}year} \PY{o}{*} \PY{n}{np}\PY{o}{.}\PY{n}{pi} \PY{o}{*} \PY{n}{c}\PY{o}{.}\PY{n}{G} \PY{o}{*} \PY{n}{c}\PY{o}{.}\PY{n}{m\PYZus{}p} \PY{o}{/} \PY{p}{(}\PY{n}{c}\PY{o}{.}\PY{n}{c} \PY{o}{*} \PY{n}{c}\PY{o}{.}\PY{n}{sigma\PYZus{}T}\PY{p}{)}
\PY{n}{res\PYZus{}factor}
\end{Verbatim}
\end{tcolorbox}

\prompt{Out}{outcolor}{17}{}

    $2.2182837 \times 10^{-9} \; \mathrm{\frac{1}{yr}}$

    Using the above calculations

\begin{equation}
L_{edd} = 2.22\times10^{-9} Mc^2 yr^{-1}$
\end{equation}

We can also write a mass accretion rate

\begin{equation}
\dot{M}_{edd} = \frac{L_{edd}}{\epsilon c^2} = \frac{2.22\times10^{-9} M}{\epsilon} yr^{-1}
\end{equation}

We are given \(\epsilon = 0.1\) in the problem giving an equation

\begin{equation}
\dot{M}_{edd} = 2.22\times10^{-8} M yr^{-1}
\end{equation}

Integrating the mass from \(M_0 \rightarrow eM_0\) and from
\(t=0 \rightarrow t=\tau\) for an e-folding time \(\tau\) gives

\begin{equation}
\int_{M_0}^{e M_0} \dot{M}_{edd} dM = 2.22\times10^{-8}  yr^{-1} \int^{\tau}_{0} M dt
\end{equation}

Assuming an exponential growth from \(M_0 \rightarrow eM_0\) then we
have a functional form \(M = M_0 e^{t/\tau}\)

\begin{equation}
M_0 (e-1) = 2.22\times10^{-8}  M_0 yr^{-1} \int^{\tau}_{0} e^{t/\tau} dt
\end{equation}

\begin{equation}
M_0 (e-1) = 2.22\times10^{-8} M_0 yr^{-1} \tau (e - 1)
\end{equation}

Simplifying gives a timescale

\begin{equation}
\tau = \frac{yr}{2.22\times10^{-8}} = 4.5\times10^7 yr
\end{equation}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{16}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZob{}}\PY{p}{(}\PY{l+m+mi}{1}\PY{o}{/}\PY{l+m+mf}{2.22e\PYZhy{}8}\PY{p}{)}\PY{l+s+si}{:}\PY{l+s+s2}{.2e}\PY{l+s+si}{\PYZcb{}}\PY{l+s+s2}{ yr}\PY{l+s+s2}{\PYZdq{}}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{16}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
'4.50e+07 yr'
\end{Verbatim}
\end{tcolorbox}

    \subsection{Problem 4}\label{problem-4}

Closed Box Enrichment In a scenario where stars are made from gas that
is initially free of metals, so Z(t=0)=0, what is the mean metal
abundance of stars once all of the gas is gone?

\textbf{My Solution}

The average stellar metalicity is defined as

\begin{equation}
<z_s> = \frac{M_z}{M_s}
\end{equation}

Where \(M_z\) is the mass of the metals and \(M_s\) is the stellar mass.
At some time t, when \(M_g = 0\) (ie all the gas was used up to make
stars), we know that \(M_{tot} = M_s\) giving

\begin{equation}
<z_s> = \frac{M_z}{M_s}
\end{equation}

For an outer shell on the star \(\delta M_s\),

\begin{equation}
M_z = \int_0^t z(t) \delta M_s
\end{equation}

Changing variables to \(M_g\)

\begin{equation}
M_z = -\int_{M_{tot}}^0 z(M_g) \delta M_g
\end{equation}

Using the equation for \(z(M_g)\) derived in class gives an integral

\begin{equation}
M_z = P \int_{M_{tot}}^0 \ln\left(\frac{M_g}{M_{tot}}\right) \delta M_g
\end{equation}

\begin{equation}
M_z = P \left[ M_g \left( \ln\frac{M_g}{M_{tot}} - 1 \right) \right]^0_{M_{tot}}
\end{equation}

\begin{equation}
M_z = P \left[ 0 + M_g\right]
\end{equation}

\begin{equation}
M_z = P ~ M_g 
\end{equation}

Substituting this back into the equation for \(<z_s>\) gives

\begin{equation}
\boxed{<z_s> = P}
\end{equation}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{ }{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]

\end{Verbatim}
\end{tcolorbox}

    % Add a bibliography block to the postdoc

\end{document}